Simulation Tools for Small Area Estimation: Introducing the \proglang{R}-package \proglang{saeSim}

### 2015-01-27 saeSim.tex
saeSim proofread

\begin{document}
%
%% include your article here, just as usual
%
% R CMD Sweave %.Rnw
% Loading (and installing) necessary packages.
%
\section{Introduction}
The demand for reliable small area statistics from sample surveys has substantially grown over the last decades due to their use in public and private sectors. In this paper we present a framework for simulation studies \red{within} the field of small area estimation. This tool might be useful for the prospective researcher or data analyst to provide reproducible research.

@@

Using this data as population, we repeatedly draw samples from it. Then we predict the domain means by using a direct estimator and a unit-level model. The sampling design is to draw a 10 per cent sample from each region with simple random sampling. For each region the direct estimator for income and the EBLUP under the BHF model is computed. Although the data offers some more information, we only use \proglang{gender}, \proglang{age} and \proglang{agesq} as covariates. The function \proglang{eblupBHF} from the package \proglang{sae} is an implementation of the BHF estimator. \red{This function takes three data }objects and returns the domain predictions. The three objects are the sampled data, the population means of the auxiliary variables and the population sizes in each domain.

\red{Before we begin to construct the simulation setup, we store these data frames as attributes to the population data. This allows us to process meta data alongside the main data frame. It is important to note that not all functions for manipulating data frames in \proglang{R} preserve these attributes. Users of \proglang{saeSim} have to keep this in mind when they implement new functions. Defining the interfaces between components differently is one possibility to avoid the usage of attributes. This can be done, for example, by using generic vectors or S4 classes instead of data frames. Yet this will add complexity to the process of data manipulation underlying the package which we try to avoid by following the paradigm: \textit{data frame in, data frame out}. Thus all functions in \proglang{saeSim} preserve the attributes of the main data frame.}

@@

\section{Outlook}
\label{sec:outlook}
From our perspective, there is a need for sharing tools for data generation \red{and simulation among the scientific community} to guarantee the reproducibility of research. \proglang{saeSim} may provide an adequate framework for pursuing this aim in the field of small area estimation. By defining the steps of a simulation we may promote a reasonable way to communicate results in academic articles and during research.

@@

The package provides more features \red{than are} introduced in this article. \red{One may mention in this context the support of outlier contaminated data. Currently only representative outliers} (outlying observations in the population) are supported \citep[cf.][]{Cha86}. However, we plan to extend this feature to non-representative outliers (outliers are part of the sample but not the population, e.g., incorrectly recorded values). Furthermore, an interface to random number generators in \proglang{R} is available. The user can also generate group effects as needed in mixed models. As the response is created by an \proglang{R} expression, any form of non-linearity in the relationship between response and auxiliary variables as well as error components can be modelled.

>>>>

\begin{document}
%
%% include your article here, just as usual
%
% R CMD Sweave %.Rnw
% Loading (and installing) necessary packages.
%
\section{Introduction}
The demand for reliable small area statistics from sample surveys has substantially grown over the last decades due to their use in public and private sectors. In this paper we present a framework for simulation studies within the field of small area estimation. This tool might be useful for the prospective researcher or data analyst to provide reproducible research.

@@

Using this data as population, we repeatedly draw samples from it. Then we predict the domain means by using a direct estimator and a unit-level model. The sampling design is to draw a 10 per cent sample from each region with simple random sampling. For each region the direct estimator for income and the EBLUP under the BHF model is computed. Although the data offers some more information, we only use \proglang{gender}, \proglang{age} and \proglang{agesq} as covariates. The function \proglang{eblupBHF} from the package \proglang{sae} is an implementation of the BHF estimator. This function takes three data objects and returns the domain predictions. The three objects are the sampled data, the population means of the auxiliary variables and the population sizes in each domain.

Before we begin to construct the simulation setup, we store these data frames as attributes to the population data. This allows us to process meta data alongside the main data frame. It is important to note that not all functions for manipulating data frames in \proglang{R} preserve these attributes. Users of \proglang{saeSim} have to keep this in mind when they implement new functions. Defining the interfaces between components differently is one possibility to avoid the usage of attributes. This can be done, for example, by using generic vectors or S4 classes instead of data frames. However this will add complexity to the process of data manipulation underlying the package which we try to avoid by following the paradigm: \textit{data frame in, data frame out}. Thus all functions in \proglang{saeSim} preserve the attributes of the main data frame.

@@

\section{Outlook}
\label{sec:outlook}
From our perspective, there is a need for sharing tools for data generation and simulation among the scientific community in order to guarantee the reproducibility of research. \proglang{saeSim} may provide an adequate framework for pursuing this aim in the field of small area estimation. By defining the steps of a simulation we may promote a reasonable way to communicate results in academic articles and during research.

@@

The package provides more features than are introduced in this article. One may mention in this context the support of outlier contaminated data. Currently only representative outliers (outlying observations in the population) are supported \citep[cf.][]{Cha86}. However, we plan to extend this feature to non-representative outliers (outliers are part of the sample but not the population, e.g., incorrectly recorded values). Furthermore, an interface to random number generators in \proglang{R} is available. The user can also generate group effects as needed in mixed models. As the response is created by an \proglang{R} expression, any form of non-linearity in the relationship between response and auxiliary variables as well as error components can be modelled.
